---
title: "Analysis"
output: html_document
date: "`r Sys.Date()`"
---
Descriptive statistics
explore data

```{r setup, include=FALSE}
str(data)
summary(data)
mean(data$sbp, na.rm = TRUE)
median(data$bmi,na.rm = TRUE)
```
Mode,quantiles, range, calculation
```{r setup, include=FALSE}
library(DescTools)
Mode(data$bmi_grp4,na.rm=TRUE)
quantile(data$bmi, probs = 0.25, na.rm = TRUE)
IQR(data$sbp, na.rm = TRUE)
range(data$sbp, na.rm = TRUE)
sd(data$sbp, na.rm = TRUE)
var(data$sbp, na.rm = TRUE)
```
Representation graphique
```{r setup, include=FALSE}
library(tidyverse)
library(RColorBrewer)
bmi.counts<-table(data$bmi)
barplot(bmi.counts, xlab = "bmi",ylab = "populaion",main = "Represenation", col=heat.colors(56))
hist(data$sbp,xlab = "xxx",ylab = "yyy",main = "Representation",col = heat.colors(67))
```


Grouping continuous data
```{r setup, include=FALSE}
data$bmi_group<-NA
data$bmi_group [data$bmi<18.5]<-0

data$bmi_group [data$bmi >=18.5 & data$bmi<25]<-1

data$bmi_group [data$bmi >=25 & data$bmi <30]<-2

data$bmi_group [data$bmi >=30]<-3
data$bmi_group <- factor (data$bmi_group, labels=c("<18.5", "18.5-24.9", "25-29.9", ">30"))
summary(data$bmi_group)
```

Standard error (SE) can be defined as the standard deviation of the sampling distribution. Standard Deviation (SD) and sample size (n) can be used to calculate SE as per the following formula:

LaTeX: SE=\frac{SD}{\sqrt{n}}

```{r setup, include=FALSE}
SBP.n <- sum(!is.na(data$sbp))
SBP.sd<-sd(data$sbp, na.rm = TRUE)
SBP.sd<-round(SBP.sd,3)
SBP.se<-SBP.sd/sqrt(SBP.n)
SBP.se<-round(SBP.se, 3)
```
Z0.975 is the 97.5% percentile of the standard normal distribution (~1.96 SD from the mean). We can determine this precisely in R using qnorm(0.975), as previously demonstrated.

Knowing this formula, and armed with both our previous calculations, and knowledge of how to calculate the mean, we can calculate the 95% Cis for SBP as follows.
```{r setup, include=FALSE}
Za <- qnorm (0.975)

SBP.mean <- mean(data$sbp, na.rm=TRUE)

ci.Z_SBP <- c(SBP.mean-(SBP.sd/sqrt(SBP.n)*Za), SBP.mean + (SBP.sd/sqrt(SBP.n)*Za))

ci.Z_SBP
```

Hypothesis testing

```{r setup, include=FALSE}
data$bmibinary <- NA
data$bmibinary[data$bmi <25] = 1
data$bmibinary[data$bmi >= 25] = 2
table (data$bmibinary)
t.test(sbp~bmibinary, data=data)
```

Form of data                Transformation

Slightly right-skewed       Square root

Moderately right-skewed     Logarithmic

Very right-skewed           Reciprocal

Left-skewed                 Square

Counts                      Square root

Proportions                 Arcsine square root

```{r setup, include=FALSE}
hist(data$hdlc,xlab = "hdlc" , ylab = "population",main = "hdlc of the whitehall people", col = heat.colors(43))
data$variable_log <- log(data$hdlc)
```

POWER & SAMPLE SIZE

Power calculations for a difference in means

cohen's = d = mean's diff/SD 
```{r setup, include=FALSE}
install.packages("pwr")
library("pwr")
power1<-pwr.t.test(d=d, power=0.9, sig.level =0.05 )
power2<-pwr.t.test(n=190, d=d, sig.level =0.05 )
```


Power Calculations for a difference in proportions
### alpha = sig.level option and is equal to 0.05
### power = 0.80
### p1 = 0.10
### p2 = 0.15
```{r setup, include=FALSE}
power4<-pwr.2p.test(h=ES.h(p1=0.1, p2=0.15), sig.level=0.05, power=0.9)
plot(power4)
```

Sample size calculation for testing a hypothesis (Clinical trials or clinical interventional studies)

```{r setup, include=FALSE}
power6<-pwr.2p.test(h=ES.h(p1=0.25, p2=0.65), power=0.8, sig.level=0.05)
```


Sample size calculations for cross-sectional studies (or surveys)

```{r setup, include=FALSE}
power8<-pwr.p.test(h=ES.h(p1=0.3, p2=0.4), power=0.8, sig.level=0.05)
```

Sample size calculations for case-control studies
p2 = (OR*p1 )/(1 +p1 (OR-1))
```{r setup, include=FALSE}
#Calcul par simple formula
```


Correlation and Scatterplots

```{r setup, include=FALSE}
plot(data$bmi,data$sbp, xlab = "BMI (kg/m2)", ylab="Systolic blood pressure (mm Hg)", cex=0.8)
cor(data$bmi,data$sbp,use="complete.obs")
```

Differences Between Means (ANOVA 1)
We use aov() to perform ANOVA, and we get a summary of the ANOVA table using summary().
aov() can be used in two ways as follows:
fit3 <- aov(y ~ x, data=my.data)
fit3 <- aov(my.data$y ~ my.data$x)

```{r setup, include=FALSE}
data$bmi_fact<-factor(data$bmi_grp4)
fit3<-aov(sbp~bmi_fact, data=data)
summary(fit3)
```

Univariable Linear Regression

```{r setup, include=FALSE}
data$bmi_fact<-factor(data$bmi_grp4)
fit3<-aov(sbp~bmi_fact, data=data)
summary(fit3)
```

Univariable Linear Regression


First, we need to tell R to treat our exposure as a categorical variable:
We can use is.factor() to check that the new created variable is indeed a factor variable.
Then, we use table() to count the number of participants in each BMI category.
We can use aggregate() to compute summary statistics for SBP by BMI.
```{r setup, include=FALSE}
data$bmi_fact<-factor(data$bmi_grp4)
is.factor(data$bmi_fact)
table(data$bmi_fact)
aggregate(data$sbp, list(data$bmi_fact), FUN=mean, na.rm=TRUE)
data<-within(data,bmi_fact<-relevel(bmi_fact,ref=2))
fit2 <- lm(sbp ~ bmi_fact, data=data)
summary(fit2)
confint(fit2, level=0.95)

```


Multivariable Linear Regression
my.fit <- lm(Y ~ X+ X2 + X3, data = my.data)

```{r setup, include=FALSE}
fit4 <- lm(sbp ~ bmi_fact+ldlc, data=data)
summary(fit4)
fit5 <- lm(sbp ~ bmi_fact+ldlc+currsmoker, data=data)
summary(fit5)
```

Model Selection and F-Tests

```{r setup, include=FALSE}
install.packages("car")
library(car)
linearHypothesis(fit5, c("bmi_fact1=0", "bmi_fact3=0", "bmi_fact4=0"))

```

Regression Diagnostics

```{r setup, include=FALSE}
hist(resid(fit5), main="Histogram of the residuals", xlab="Residuals")
sum(resid(fit5))
plot(resid(fit5) ~ fitted(fit5), main="The residuals versus the fitted values", xlab="Fitted values", ylab="Residuals", cex=0.8)
abline(0,0, col=2)
```



